---
title: "Synthetic Control in robustcause"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Synthetic Control in robustcause}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Overview

`robustcause::fit_sc()` exposes both standard synthetic control and a robust
MM synthetic-control estimator through a single R interface. The robust
variant keeps donor weights on the simplex but downweights contaminated
pre-treatment periods during fitting.

The robust MM estimator helps when a few pre-treatment periods are distorted by
measurement error, transient shocks, or heavy-tailed noise. It does not solve
fundamental donor-pool mismatch or lack of overlap.

## Matrix interface

```{r}
library(robustcause)

outcomes <- cbind(
  treated = c(1, 2, 20, 10, 11, 12),
  d1 = c(1, 2, 2.8, 3, 3.2, 3.3),
  d2 = c(0.9, 1.9, 3.1, 3.5, 3.9, 4.1),
  d3 = c(1.2, 2.1, 2.9, 3.2, 3.4, 3.6)
)

fit_mm <- fit_sc(
  outcomes = outcomes,
  treated_unit = "treated",
  treatment_start = 4,
  method = "mm",
  run_placebos = TRUE
)

summary(fit_mm)
coef(fit_mm)
head(as.data.frame(fit_mm))
```

The fitted object supports standard extraction helpers:

```{r}
fitted(fit_mm)
residuals(fit_mm)
predict(fit_mm, type = "gap", period = "post")
```

## Panel-data interface

```{r}
panel <- data.frame(
  unit = rep(c("treated", "d1", "d2", "d3"), each = 6),
  time = rep(1:6, times = 4),
  outcome = c(
    1, 2, 20, 10, 11, 12,
    1, 2, 2.8, 3, 3.2, 3.3,
    0.9, 1.9, 3.1, 3.5, 3.9, 4.1,
    1.2, 2.1, 2.9, 3.2, 3.4, 3.6
  ),
  predictor_a = c(
    5, 5, 5, 9, 9, 9,
    5, 5, 5, 6, 6, 6,
    4, 4, 4, 7, 7, 7,
    6, 6, 6, 8, 8, 8
  ),
  predictor_b = c(
    2, 2, 2, 4, 4, 4,
    2, 2, 2, 3, 3, 3,
    1, 1, 1, 3, 3, 3,
    2, 2, 2, 3, 3, 3
  )
)

fit_panel <- fit_sc(
  data = panel,
  unit = "unit",
  time = "time",
  outcome = "outcome",
  treated_unit = "treated",
  treatment_start = 4,
  predictors = c("predictor_a", "predictor_b"),
  predictor_weights = c(2, 1),
  method = "standard",
  run_placebos = FALSE
)

summary(fit_panel)
```

## Diagnostics and plots

`fit_sc()` stores the full treated, synthetic, and gap path in a convenient
data-frame form.

```{r}
tail(as.data.frame(fit_mm))
```

Base plotting methods are available for the gap path, donor weights, and
placebo post-treatment gaps.

```{r, fig.height=4, fig.width=6}
plot(fit_mm)
plot(fit_mm, type = "weights")
plot(fit_mm, type = "placebo_gaps")
```

## Missing values

The wrapper currently handles missingness through `na.action`:

- `"fail"` stops when incomplete rows or periods are present
- `"omit"` drops incomplete rows for panel data or incomplete periods for
  matrix input before fitting

The compiled solver itself still expects finite matrices after preprocessing.
